Armour can be used to reduce injury level (not victory level, which only Strength can reduce.) Wearing lots of armour will also add weight which negatively affects a player’s combat roll.

Armour levels are set by the type of armour the character is wearing, described below. In the case of shields, the armour rating stacks with worn armour. A character wearing a full plate (4 armour) and using a large shield (2 armour) will have an armour rating of 6. This is the highest armour rating in the game, but will also remove -3D6 from their combat roll

PLEASE NOTE: Injury reduction from armour is not applicable while in Close Combat (See Combat Modifiers) %cite!

\begin{center}
    \begin{xltabular}{\textwidth}{|l l X|} 
        \hline 
        \textbf{Armour Level} & \textbf{Combat Roll} & \textbf{Examples} \\ 
        \hline
        \textbf{1} & -0D6 & 
        $\bullet$ Leather Armour \newline %cant find a way to use \itemize and have it look good so we'll do bullets manually
        $\bullet$ Gambeson (padded) \newline 
        $\bullet$ Chain Shirt (hauberk) \newline 
        $\bullet$ Small Shield \vspace{1em} \\
        \textbf{2} & -1D6 & 
        $\bullet$ Breastplate \newline 
        $\bullet$ Full Chainmail \newline 
        $\bullet$ Large Shield \vspace{1em} \\
        \textbf{3} & -2D6 & 
        $\bullet$ Full Plate armour \vspace{1em} \\
        \textbf{4} & -3D6 & 
        $\bullet$ Full Plate armour + Small Shield \vspace{1em} \\
        \textbf{5} & -4D6 & 
        $\bullet$ Full Plate armour + Large Shield \vspace{1em} \\
        \hline
    \end{xltabular}
\end{center}

A character can reduce injury levels by the equivalent amount of levels they have in armour, per combat encounter. When another fight begins (another combat encounter) the character’s pre-set injury reduction from their armour level is reset.

Example:

\begin{displayquote}
    Victoria is a noble knight and is attacked unawares by a backstabbing rogue.

    The rogue achieves a level 3 victory and chooses to inflict a serious injury. Victoria is wearing full plate armour and can reduce up to 4 points of injury throughout this conflict.

    Reducing this serious injury by 3 points times will negate it; one to reduce it to a medium injury, a second to reduce it further to a minor injury, and another to remove the injury entirely.

    Victoria chooses to do so, which will leave her one last injury reduction from her armour for that combat encounter, but enable to fight without such a disadvantage. She will however, be fighting with -2D6 to her combat roll due to her heavy armour.
\end{displayquote}

When achieving a victory in combat you can choose to inflict an injury – but if your opponent is wearing armour choosing to inflict an injury will be more easily countered through their injury reduction. However, other actions chosen through a victory, such as staggering, disarming or tripping an armoured opponent, cannot be negated with armour (see Combat Modifiers, CH4, PT7.)

For a more sophisticated approach to the armour rules within Cogent Roleplay, please see Chapter 6: Hardcore Combat

\section{Endurance and Armour} \label{sec:endurance_and_armour}

A character with high endurance is considered much more capable than average of functioning with heavy armour on. As such, for every point in the Endurance skill (NOT including the base 3D6 or any strength modifiers) the negative to the combat roll is reduced. A character with 1 point in endurance will not receive -1D6 to their combat roll for carrying a small shield or wearing a padded gambeson. A character with 4 points in endurance can wear full-plate mail and carry a large shield and receive the benefits of damage reduction that come with armour, without any of the negative combat modifiers.