When a character's life is on the line, every asset they possess is brought to bear in order to win. Thus every attribute; strength, reflex and intelligence, can be added to the combat roll.

Combat Dice Pools are calculated by adding:
\begin{itemize}[label=+]
    \item[~] Base \textbf{3D6}
    \item \textbf{1D6} for each Attribute point (\textbf{STR}/\textbf{REF}/\textbf{INT})
    \item \textbf{1D6} for each point in any \textbf{Combat Skill} (Relevant to what the character is wielding or doing)
    \item Any applicable \textbf{weapon bonus}
\end{itemize}

An example follows:
\begin{displayquote}
Sir Terrik the Knight has challenged another knight to a duel. He has the vocation "Knight" and nested under that vocation he has taken the combat skill "Medium Weapons" to showcase his prowess with the sword. Terrik's Dice Pool is calculated thus:
Base of 3D6, plus +1D6 from STR and +1D6 from INT
2D6 from his two skill points invested in the "Medium Weapons" Skill    (which can be applied as he is wielding a medium sword)
+1D6 from his arming sword (the bonus for a medium weapon)
His total Combat Dice Pool is 8D6.
\end{displayquote}

\section{Victory Levels}\label{sec:victory_levels}

A successful combat roll results in a level of \textbf{victory} against your opponent(s). Your combat roll stands as both an \textbf{attack} and \textbf{defence}. It represents your character’s overall combative performance for that round and not necessarily a single ‘swing’ of their weapon.

Each character’s combat roll is compared to that of the character(s) or non-player character(s) they oppose. Every \textbf{win} a character has over another’s roll, means a \textbf{level of victory}, from which they may choose an outcome.

SEE CHEAT SHEET REFERENCE: CHSH - Victory Levels

\subsection{Level 1 Victory} \label{subsec:level_1_victory}

\begin{itemize}
    \item Minor Injury, the opponent receives \textbf{-1D6} to all physical-based dice checks but will naturally heal over the next few days, or faster if treated actively. This can be negated by strength or armour.
    \item Stagger (\textbf{-2D6} for one round)
    \item Enter into \textbf{close combat} (not available to ranged combat rolls)
\end{itemize}

\subsection{Level 2 Victory} \label{subsec:level_2_victory}

\begin{itemize}
    \item Medium injury \textbf{-2D6} to all physical-based dice checks and must be treated relatively soon, otherwise further penalties may apply (infection etc.) – this can be reduced by strength or armour to a lower injury or fully negated.
    \item Disarm
    \item Any lower outcome
\end{itemize}

\subsection{Level 3 Victory} \label{subsec:level_3_victory}

\begin{itemize}
    \item Serious injury \textbf{-3D6} to all physical-based dice checks and the character will die in time if the wound is untreated. This can be reduced by strength or armour to a lower injury or fully negated.
    \item Trip
    \item Any lower outcome
\end{itemize}

\subsection{Level 4 Victory} \label{subsec:level_4_victory}

\begin{itemize}
    \item Fatal injury \textbf{-4D6} to all physical-based dice checks, incapacitated, and will die very soon if untreated. This can be reduced by strength or armour to a lower injury or fully negated.
    \item Knockout Blow – Render your opponent unconscious, and apply a Medium Injury to the opponent in addition.
    \item Dismember an arm or leg, which also gives a serious injury to the opponent. (not available to ranged combat rolls unless magic or explosive projectiles are involved)
    \item Any lower outcome
\end{itemize}

\subsection{Level 5 Victory} \label{subsec:level_5_victory}

\begin{itemize}
    \item Deathblow. An \textbf{instant kill} unless the character has high strength and or is wearing armour and uses them to reduce either the level of victory or injury.
    \item Any lower outcome
\end{itemize}

Injury levels do not stack in fatality but do stack in dice penalty – meaning two medium injuries do not equate to a fatal injury where the character will die immediately if untreated, but the character does receive \textbf{-2D6} to all physical dice checks TWICE, meaning they will have \textbf{-4D6} in total.

\section{Defence Rolls} \label{sec:defence_rolls}

A Defence Roll is a variant of the Combat Roll. It is used when a character is either in a position where they cannot perform any type of Combat Manoeuvre to their opponent (e.g. too far away from an opponent using a ranged weapon against them while only armed with a melee weapon), or they want a greater chance to survive as their standard Combat Roll is too low (such as fleeing after receiving an injury.)

A Defence Roll is calculated in the same way as the Combat Roll but adds an additional \textbf{+2D6} to roll with. No matter the result of the Defence Roll, the character cannot achieve a level of victory because the character was purely focused on protecting themselves, and not on attacking.