Each weapon inherently provides advantages over other weapons. The following is a guide and can be altered as the Narrator sees fit.

SEE CHEAT SHEET REFERENCE: CHSH - Weapon Modifiers

There are eight types of \textbf{Weapon Group}, each affecting your combat role. They are as follows:

\begin{center}
    \begin{xltabular}{\textwidth}{|l X X|} 
        \hline 
        \textbf{Weapon Group} & \textbf{Examples} & \textbf{Combat Roll}\\ 
        \hline
        Ranged & & 
                $\bullet$ +2D6 (Ranged Target) \newline
                $\bullet$ -3D6 (Melee Target) \vspace{1em} \\
        Unarmed & 
            $\bullet$ Fist \vspace{1em} & 
                $\bullet$ -1D6 (Ungloved) \newline
                $\bullet$ +0D6 (Gloved) \vspace{1em} \\
        Small Weapon & 
            $\bullet$ Dagger \newline
            $\bullet$ Hatchet \newline
            $\bullet$ Shortsword \vspace{1em} & 
                $\bullet$ +0D6 \vspace{1em} \\
        Medium Weapon & 
            $\bullet$ Sword  \newline
            $\bullet$ Battleaxe \newline
            $\bullet$ Warhammer \vspace{1em} & 
                $\bullet$ +1D6 \vspace{1em} \\
        Large Weapon & 
            $\bullet$ Longsword \newline
            $\bullet$ Dane axe \vspace{1em} &
                $\bullet$ +2D6 \vspace{1em} \\
        Reach Weapon & 
            $\bullet$ Polearm \newline
            $\bullet$ Spear \newline
            $\bullet$ Greatsword \vspace{1em} &
                $\bullet$ +3D6 \vspace{1em} \\
        Hand Ballistic & 
            $\bullet$ Handgun \newline
            $\bullet$ SMG \vspace{1em} &
                $\bullet$ +2D6 (Ranged Target) \newline
                $\bullet$ -3D6 (Melee Target) \vspace{1em} \\
        Shoulder Stock Ballistic & 
            $\bullet$ Hunting Rifle \newline
            $\bullet$ Shotgun \newline
            $\bullet$ Assault Rifle \vspace{1em} &
                $\bullet$ +4D6 (Ranged Target) \newline
                $\bullet$ -4D6 (Melee Target) \vspace{1em} \\
        Heavy Ballistic & 
            $\bullet$ Bazookas \newline
            $\bullet$ Heavy Machine Gun \newline
            $\bullet$ Minigun \vspace{1em} &
                $\bullet$ +6D6 (Ranged Target) \vspace{1em} \\
        \hline
    \end{xltabular}
\end{center}

Some of these weapons also have other limitations/abilities:

\begin{displayquote}
    \subsubsection{Unarmed - Disarm Bonus}

    Can be used in close combat and receives a bonus in disarming actions. A player must be in close combat to disarm another player while being unarmed.

    Once in close combat, if declaring to attempt a disarm, the player receives +4D6 to their combat roll, but the roll cannot be used to inflict any other injury victory (this means they cannot delegate their attack to inflict an injury or execute a trip regardless of how high the attack roll is.)

    \subsubsection{Small Weapons}

    Small weapons can be used in close combat.

    \subsubsection{Reach Weapons}

    Reach is a significant advantage in combat, and is reflected in the higher weapon bonuses – but the Narrator is free to negate all dice bonuses if they feel the reach weapon would not work as effectively due to narrow confines such as an alleyway or hallway.

    Medium weapons and smaller can be used to enter into close combat against any character using reach weapons (see close combat under combat modifiers.) %cite!

    \subsubsection{Heavy Ballistic}

    Heavy ballistic weapons can't be used in melee combat.
\end{displayquote}

For a more sophisticated approach to the weapon modifiers and ranged combat rules within Cogent Roleplay, please see chapter six: Hardcore rules. %cite

SEE CHEAT SHEET REFERENCE: CHSH - Ballistic Modifier.