o far untouched in the text of this manual, the Destiny Roll stands alone as an extremely powerful and dynamic storytelling tool. A destiny roll is made by a 20-sided dice, or D20, and is made by the Narrator, or assigned by the Narrator to be made by a player.

This roll can be applied to endless scenarios and is used to allow the Narrator to let fate – the roll of a dice – decide what direction the story will move on in, be that for or against the player's goals and desires.

The higher the roll is, and the closer the roll is to 20, the more advantageous the result is to the player(s). Adversely, the lower the roll and the closer it is to 1, the worse the disadvantage. Furthermore, rolling a 20 is considered a critical success, and may allow the Narrator to assign exceptionally advantageous circumstances to the party or player. Likewise, rolling a 1 is always to be considered a critical failure, and will result in all kinds of things the players won’t want. Whether they be revealed as immediate or delayed consequences is up to the Narrator.

These rolls may be used in situations of great intensity, to help the Narrator guide the path of the action. They may be used to help break up the monotony at random times to see if something good or bad encounters the player, or even what the weather is (see CH7, “Random Encounter Tables”), and they may also be used to determine the intensity of consequence of a players choice.

Here are a few simple examples of the Destiny Roll in action:

\begin{center}
    \begin{tabular}{|p{0.22\columnwidth}|p{0.22\columnwidth} p{0.22\columnwidth} p{0.22\columnwidth}|} 
        \hline 
        \textbf{Scenario} & \textbf{Destiny Roll < 5} & \textbf{Destiny Roll 6 - 14} & \textbf{Destiny Roll > 15} \\ 
        \hline
        A battle has been waged for several rounds, and the enemy is about to unveil a new threat… & 
        A colossal dragon, loosely chained and ready to attack anything in sight! & 
        A squad of angry soldiers led by a menacing captain. & 
        A siege weapon preparing to fire, but it seems to be constructed late and might fall apart. \\
        The players enter an alley and are greeted by… &
        A gang of thieves, with weapons drawn. &
        The town drunk, who looks confused. &
        An old acquaintance with a skill-set the party needs who is keen for a drink and a catch-up. \\
        You fall through the floor into the room below into… &
        A dungeon, with no apparent means of escape… &
        A courthouse, whereupon several officials gasp in surprise! &
        A small religious ceremony, where the faithful are praying for the chosen one to appear. \\
        \hline
    \end{tabular}
\end{center}

