Please note that even though this system is, as of yet, unfinished, it can still be used as the system works off a CL basis. All the player has to do is describe what they wish to do with their spell and the narrator will decide if the spell applies to the action the player is requesting to perform, and then set the CL they must match in order to cast it.

This simple pre-built magic system is given to demonstrate how magic can be incorporated into the Cogent System. It is made to reflect classic fantasy and can be used in your settings or used as an example for one that you make for yourself.

We do recommend that you at least use the individual spells as models for other superpowers or magic systems that you might want to incorporate. For instance the rules and CLs for the Telepathy spell as outlined below work perfectly for a telepathy superpower. Feel free to take and rename what you wish.

It is also made to be perfectly balanced (to the best of our ability) so magic users won’t be overpowered or disadvantaged.

\section{Introduction to the Classical Fantasy Magic System}

Each spell is an individual proficiency that a player can spend up to two skill points in. They are decently versatile and there’s no limit to the number of times they can be used.

Each spell falls into one or more of the schools or disciplines of magic (wizardry, sorcery, etc) and a player must select a school before they choose any spells to know. It does not cost anything to choose a school of magic but once a school is chosen the player can only select spells from that specific school. This is because each school of magic draws its power from different sources (really it’s for additional game balance and to remain true to the classic fantasy tropes.)

Of course, if the narrator feels that a specific skill fits more with a different school of magic they are more than free to mix and match as they see fit.

The player can choose how to use the spell so long as it falls within the spells guidelines and the Narrator will assign a CL (based on the standard Challenge Level difficulties, 1 common task, 2 uncommon, 3 specialized, etc) that must be matched for the spell to be cast. This means there’s an element of failure in using magic. Failure according to this magic system never means the spell was not cast, it means the spell was miscast, meaning something other than what the player wanted to happen occurred. The level of failure from the CL will ascribe how great a miscast it was and the narrator will fill in the rest.

Some spells summon elementals according to the summoning CL table below.

Please note that the number of wins exceeding the CL will determine how long the elemental will remain in the physical plane.

\begin{itemize}
    \item Matching the CL = three rounds
    \item +1 over the CL = ten rounds
    \item +2 over the CL = one hour
    \item +3 over the CL = six hours
    \item +4 over the CL = a day
    \item +5 over the CL = will stay with sorcerer until dismissed
\end{itemize}

Rolling under the CL will cause the elemental to attack the sorcerer and it will remain in the world for the equivalent amount of losses under the CL as wins over the CL determines.

\begin{center}
    \begin{tabular}{|l l l|} 
      \hline
      \textbf{CL} & \textbf{Summon} & \textbf{Effect} \\ 
      \hline
      CL~1 & Little Sprite & 3d6 combat roll \\ 
      CL~2 & Sprite & 4d6 combat roll \\ 
      CL~3 & Greater Sprite & 5d6 combat roll \\ 
      CL~4 & Lesser Elemental & 7d6 combat roll \\ 
      CL~5 & True Elemental & 8d6 combat roll \\ 
      CL~6 & Greater Elemental & 9d6 combat roll \\ 
      CL~7 & Lord Elemental & 10d6 combat roll \\ 
      CL~8 & Elder Elemental & 11d6 combat roll \\ 
      \hline
    \end{tabular}
\end{center}

Please note that elemental combat rolls can strike multiple opponents

Elementals can only be damaged by other elemental based attacks (they are immune to their own element), magic attacks and receive double damage from attacks by their opposing element. When elementals are struck by a damage type that can harm them, they receive damage in the same way as any other creature and character, meaning an injury severity from combat manoeuvres.

\section{Sorcery Spells (proficiencies)}

\subsection{Pyromancy}

This spell creates and controls fire. It can be used in all the varied and versatile ways that fire can.

This spell can also be used offensively as a base combat roll. The amount of fire that the sorcerer wants to create will determine the CL, set by the Narrator.

This spell can summon elementals.

\subsection{Hydromancy}

Creates and controls water, the amount and level of control will set the CL determined by the narrator.

This spell can summon elementals.

\subsection{Aeromancy}

Control air, fly, and create gusts of wind.

Aeromancy cannot be used as a base combat proficiency though it can be used to push on people and blow them over.

Aeromancy cannot be used to levitate things, it is too wild and strong for such delicate and controlled movement. Flying through Aeromancy is achieved by thickening the air under the magic user and pushing it past in the direction you want to fly, thus the magic user looks to be in the midst of a windstorm that is solely localized around themselves. This cannot be done to another person or object as thickening the air like this is exponentially harder the farther away from the caster.

\begin{center}
    \begin{tabular}{|c m{.8\textwidth}|} 
      \hline
      \textbf{CL} & \textbf{Summon} \\ 
      \hline
      CL~2 & Fly at standard sprinting speed. Failure does not mean you fall from the sky, but you lose control and fly in an undesired direction. \\ 
      CL~4 & Fly at speeds up to 100 km. Failure does not mean you fall from the sky, but you lose control and fly in an undesired direction. \\ 
      CL~6 & Fly at speeds up to 500 km. Failure does not mean you fall from the sky, but you lose control and fly in an undesired direction. \\
      \hline
    \end{tabular}
\end{center}

If using this spell to create a strong gust of wind to knock another player or NPC off their feet, the caster makes a conflict dice roll with their Aeromancy proficiency vs another player/NPC defence roll. Weapon bonuses cannot be used in the defence roll against this type of attack.

This spell can summon elementals.

\subsection{Terramancy}

Control and manipulate the earth by shaping it and creating earthquakes.

\section{Wizardry Spells (proficiencies)}

\subsection{Divination}

Read someone’s future, see through a focusing object (like a crystal ball or calm pool of water) to spy on people or places far away. Communicate over long distances. Prophesy, be warned of danger or traps that lay in wait ahead.

Divination can be used to assist, perception checks, or supplant perception checks if higher, and not be limited by things that are hidden. Of course the farther away things are the higher CL the Narrator will apply.

This spell can summon elementals.

\subsection{Levitation}

Any object of human size and lower doesn’t take a dice check to control unless you are trying to do one of two things:

\begin{itemize}
    \item Hold another person still who is actively trying to resist. In that case, it is a Conflict CL.
    \item Manipulate something at a distance or something heavier than normal human size and weight.
\end{itemize}

You can manipulate weapons and attack with them as if you were holding them but you do not use your Telekinesis Proficiency in the attack roll. Instead, you use whatever proficiency that weapon requires.

Using levitation to throw things as projectiles uses your levitation proficiency

The following CLs are to be used for the specified actions:

\begin{center}
    \begin{tabular}{|c m{.8\textwidth}|} 
      \hline
      \textbf{CL} & \textbf{Spell Effect} \\ 
      \hline
      No~CL & 
        $\bullet$ Manipulate something within 10 meters under human size and weight \\ \hline
      CL~3 & 
        $\bullet$ Manipulate something within 10 meters and up to car size and weight \newline
        $\bullet$ Manipulate something from 10 to 20 meters away under human size and weight \\ \hline
      CL~5 & 
        $\bullet$ Manipulate something within 10 meters and up to bus size and weight \newline
        $\bullet$ Manipulate something from 10 to 20 meters away up to car size and weight \newline
        $\bullet$ Manipulate something from 20 to 50 meters away under human size and weight \\ \hline
      CL~7 &
        $\bullet$ Manipulate something within 10 meters and up to airplane size and weight \newline
        $\bullet$ Manipulate something from 10 to 20 meters away up to bus size and weight \newline
        $\bullet$ Manipulate something from 20 to 50 meters away up to car size and weight \newline
        $\bullet$ Manipulate something from 50 to 100 meters away under human size and weight \\
      \hline
    \end{tabular}
\end{center}

This spell can summon elementals.

\subsection{Telepathy}

No dice checks are required when trying to read or communicate to a non-resisting person’s mind. A player or NPC can try and resist telepathy with an intelligence roll or even their own Telepathy proficiency. If a person is resisting you must exceed their roll to read what they are thinking, exceed their roll by 2 to probe their mind and look for specific thoughts or memories and exceed their role by 4 to inflict an injury level. Each dice beyond the four is an additional injury level inflicted.

\subsection{Counteraction}

With this spell the wizard can counter any other spell from any other type of magic. Countering a spell can do one of two things:

If the Counter Spell roll was equal to the roll of the spell it’s trying to counter, it can prevent the spell from being cast.

If the Counter Spell roll is greater than the spell it is trying to counter the wizard casting the Counter Spell can choose for the spell they are countering to be miscast (see miscast above.) The level of miscast depends on how much the wizard exceeded the spell roll of the spell they are countering.

\subsection{Self-Teleportation}

This is the ability to teleport yourself (no other objects) to desired locations. Please note that all CLs are doubled if trying to teleport to an unseen location

\begin{center}
    \begin{tabular}{|c m{.8\textwidth}|} 
      \hline
      \textbf{CL} & \textbf{Spell Effect} \\ 
      \hline
      No~CL & 
        $\bullet$ Teleport yourself to seen location \\ \hline
      CL~3 & 
        $\bullet$ Teleport yourself and something up to car size and weight to seen location within 10 meters  \newline
        $\bullet$ Teleport yourself to seen location within 20 meters \\ \hline
      CL~5 & 
        $\bullet$ Teleport yourself and something up to bus size and weight to seen location within 10 meters  \newline
        $\bullet$ Teleport yourself and something up to car size and weight to seen location within 20 meters \newline
        $\bullet$ Teleport yourself to seen location within 50 meters \\ \hline
      CL~7 &
        $\bullet$ Teleport yourself and something up to airplane size and weight to seen location within 10 meters \newline
        $\bullet$ Teleport yourself and something up to bus size and weight to seen location within 20 meters \newline
        $\bullet$ Teleport yourself and something up to car size and weight to seen location within 50 meters \newline
        $\bullet$ Teleport yourself to seen location up to 100 meters away \\
      \hline
    \end{tabular}
\end{center}

\subsection{Teleporting Others (Other-Teleportation)}

This is the ability to teleport other objects, not oneself to desired locations. You must be able to see the object to teleport it. If you are teleporting something close to you to a distant location, always use the farthest distance to determine the CL. Please note that all CLs are doubled if trying to teleport something to an unseen location

\begin{center}
    \begin{tabular}{|c m{.8\textwidth}|} 
      \hline
      \textbf{CL} & \textbf{Spell Effect} \\ 
      \hline
      CL~5 & 
        $\bullet$ Teleport something 10 meters away from its location under human size and weight \\ \hline
      CL~7 & 
        $\bullet$ Teleport something 10 meters away from its location up to car size and weight \newline
        $\bullet$ Teleport something from 10 to 20 meters away from its location under human size and weight \\ \hline
      CL~9 & 
        $\bullet$ Teleport something within 10 meters away from its location and up to bus size and weight \newline
        $\bullet$ Teleport something from 10 to 20 meters away from its location up to car size and weight \newline
        $\bullet$ Teleport something from 20 to 50 meters away from its location under human size and weight \\ \hline
      CL~12 &
        $\bullet$ Teleport something within 10 meters away from its location and up to airplane size and weight \newline
        $\bullet$ Teleport something from 10 to 20 meters away from its location up to bus size and weight \newline
        $\bullet$ Teleport something from 20 to 50 meters away from its location up to car size and weight \newline
        $\bullet$ Teleport something from 50 to 100 meters away from its location under human size and weight to seen location \\
      \hline
    \end{tabular}
\end{center}

\subsection{Prestidigitation}

\begin{itemize}
    \item Create soft light
    \item Reduce soft heat/cold
    \item Tiny illusion
    \item Create fog
    \item Create a protective shield of magic
    \item Small magical tricks
\end{itemize}

\subsection{Illusion}

Invisibility is included in illusion

\section{Witchcraft Spells (proficiencies)}

\subsection{Alteration}

Can change the composition of an element or change it into a lighter element. Making something into a lighter element will increase its size. It is possible to change something into a heavier element but requires much more power and exhausts the witch more. The heavier the change the harder it is.

\subsection{Transform Objects}

This is complex and requires a lot of power. Using this spell will usually drain the witch for a day

Chemical effects in substances can be completely reversed through alteration, such as altering a poison into a life-saving, injury-healing potion.

\subsection{Polymorph}

Transforms people and themselves into other creatures. The greater the difference from the original form, the greater the CL

\subsection{Domination}

To work this spell the Witch needs a physical thing representing their target, a doll that has a fingernail or lock of hair from the person they wish to dominate.

This spell begins with manipulating the emotions in their target, then being able to plant thoughts and ideas into their minds and ultimately taking over their body for a few rounds. The Narrator will set the CL for each different application of this spell.

\subsection{Sympathy}

This magic bonds two things and then can switch properties between the two, such as passing on the properties of one thing to the other making the other, weaker, stronger, lighter or heavier.

When two things are linked or bonded through sympathy, with a high enough CL their physical locations can be switched, no matter their location. And ultimately through sympathy, two bonded creatures can have their consciousness switched either temporarily, or permanently.

\section{Druidism Spells (proficiencies)}

\subsection{Naturmancy}

Can control non-intelligent living things, cause trees to move and vines to snare, or wood to warp.

Cause a natural element to harden or soften, wood stronger than steel, steel softer than butter (causes the thing to decay including people, or causes weapons to rust extremely fast.)

Can summon forest elementals.

\subsection{Reflection}

Can pass on properties such as injury, poison or strength to another person

Heal (not nearly as potent as holy magic but does work to an extent.)

Rotten food will become good again, old weathered wood renewed. It essentially turns back time for a specific object. This can also make a person more youthful

Causes one to sleep, be sick or feel healthy. Never permanent and cannot kill.

\subsection{Communion}

Can speak with any living thing and ask for their help. The CL will reflect the level of help being requested.

\section{Necromancy (proficiency)}

Necromancy is a spell and school unto itself.
Raise undead.
Harm and Kill with a command.
When killed can raise themselves as an undead.

\section{Holy Miracles (proficiencies)}

\subsection{Restoration}

Heal (Harms undead and any demon-possessed person or creature)
Injury
Disease
Disability
Death
Cleanse bad food or water

\subsection{Sanctification}

Increase the abilities or skills in others
Protection wards
Cast out evil
Burn corruptions with a touch

\subsection{Creation}

Create food and water
Create an object, the more complex and large, the greater the CL

\subsection{Revelation}

Works very much like the wizard’s Divination
Prophesy and foretelling (comes randomly)
Tongues (communicate with any person)
Be warned of danger ahead
Speaking with spirits, communication with the dead

\subsection{Exultation}

Walk through walls or on water or air