Often, settings for games of cogent will be diverse and unique. Games that take place in fantasy settings with wizards and witches as often as science fiction settings with hologram projectors and rocket shoes. All of these settings contain the potential for things to operate in supernatural or superhuman ways that push beyond the limitations outlaid in the Cogent system thus far.

As a general guide, there are two broad ways of achieving this.

\begin{itemize}
    \item Equipment and Items: By using equipment you effectively control the flow of these abilities through the addition of custom items. Players can obtain and use the items throughout the course of their gameplay, these could have requirements or drawbacks to use, or they may be usable with no downside once acquired. By using equipment as the source of the metahuman powers in your setting, you tie them to the commerce system or your own currency system and can moderate how accessible the most potent powers are.
    \item Vocations and Vocation Skills: Players use vocations such as “Wizard” to select Vocation Skills and Combat Skills that are suitable for your setting. It is possible that you may choose to allow magical Combat Skills to be used in other situations, or you may decide to balance the potential power of magic that players must invest skill points in the combat AND non-combat applications of a given type of magic. For example, an Elementalist with the Vocation “Earthmage” may need to take points in an “Earthshaping” Vocational Skill in order to craft things from stone using their power but must take “Earth-Fighting” as a Vocational Skill in order to use the elemental earth in combat.
    \item Vocations and Vocation Skills: Players use vocations such as “Wizard” to select Vocation Skills and Combat Skills that are suitable for your setting. It is possible that you may choose to allow magical Combat Skills to be used in other situations, or you may decide to balance the potential power of magic that players must invest skill points in the combat AND non-combat applications of a given type of magic. For example, an Elementalist with the Vocation “Earthmage” may need to take points in an “Earthshaping” Vocational Skill in order to craft things from stone using their power but must take “Earth-Fighting” as a Vocational Skill in order to use the elemental earth in combat.
\end{itemize}

These are just some ideas, and many Narrators will choose to use similar systems or blends of the two. There is no perfect solution, what matters is what works best for the story you would like to tell, and that is one of the greatest advantages of Cogent Roleplay – It is infinitely flexible!

Nevertheless, presented here to use as an example or inspiration is the same school of magic, in both hard and soft magic systems. The school of magic we will use as an example is “Druidic”

A \textbf{hard} magic system is one where the boundaries are very deliberate and precise, the exact functions you can perform are outlined in specific spells or items, and players have a clear understanding of exactly what they can cast and what it will do.

A \textbf{soft} magic system is one where the limits of what a spellcaster can do are more of guidelines, with challenge level reflecting “roughly what you could do” exerting that level of power with that school of magic. There are pros and cons to each system, and many people favour one over the other, but once again, the correct choice is the one that your gaming group wants to play!

\textit{Example of Druid Spells in a \textbf{Hard} magic system:}

\begin{center}
  \begin{tabular}{|m{0.55\textwidth} m{0.35\textwidth}|} 
      \hline 
      
      \textbf{Spell} & \textbf{Effect} \\ 
      
      \hline

      \hspace*{.8em} \textbf{Grass-shackle: CL~4} \newline
      \hspace*{.8em} \textit{Target: One other creature} \newline \newline
      Requires 2 skill points invested in “Druidic” Combat Skill. \newline \newline
      Must be holding a handful of grass gathered in moonlight, which the spell consumes. & 
        On success, a single target is bound by grass that ties itself around its feet. The bound target must fight defensively next combat round. If they do not, or they attempt to move, they are \textbf{staggered}. \\
      
      \hline

      \hspace*{.8em} \textbf{Strength of Wood: CL~4} \newline
      \hspace*{.8em} \textit{Target: Self} \newline \newline
      Requires 2 skill points invested in “Druidic Plantcraft” Vocational Skill.\newline \newline
      Must be touching a tree aged twenty or more years old. &
        On success, thick bark flows from the tree to your flesh, magically with your body until the next moonrise, where it disappears. You immediately gain \textbf{2 armour points}. This spell does not stack with itself. \\
      
      \hline

      \hspace*{.8em} \textbf{Bloom: CL~2} \newline
      \hspace*{.8em} \textit{Target: A single bud or sprout} \newline \newline
      Requires 1 skill point invested in “Druidic Plantcraft” Vocational Skill.\newline \newline
      Must be cast in the moonlight. &
        On success, the bud or sprout blooms and grows as if a week had passed. Under the effect of this spell, a flower will never wilt.\\

      \hline

      \hspace*{.8em} \textbf{Earthrot: CL6} \newline
      \hspace*{.8em} \textit{Target: One creature you are touching.} \newline \newline
      Requires 3 skill points invested in “Druidic” Combat Skill.\newline \newline
      Must be wearing the skull of a goat slaughtered on a full moon.\newline \newline
      Requires the heart of a goat, which the spell consumes.&
        On success, your target immediately suffers a \textbf{serious injury} as their flesh withers and rots. They become \textbf{infested}* with worms that burrow through the soil and enter their flesh. On a failure, you suffer a minor injury, and the goat skull splits, becoming unusable.\\
      
      \hline
  \end{tabular}
\end{center}

\textit{*Infested is not a condition that currently exists in Cogent Roleplay, but was included to show how you could also make up custom status conditions for your games and magic systems.}