These charts are mainly for use by the narrator of the Campaign and can be adapted to other genres.

With any encounter, there is still room to move. They can be small or severe in their fulfilment. With each encounter, a severity roll must accompany it to determine how extreme or mild it is. This applies to weather and monsters and is done with a D6.

These work thus:

\begin{itemize}
    \item 1-2 Mild
    \item 3-4 Average
    \item 5-6 Heavy
\end{itemize}

\section{Weather}

There are two rolls for a weather check. The first is according to the following weather charts on a 1d100 (2D10 dice.) The second is a severity check, is it drizzling or heavy rain?

The weather check should last for about two days but you can have them last for one day or a week, it’s your choice but two days represents the most realistic weather conditions. A new weather roll every day could represent turbulent weather conditions at that time in the game.

The severity check determines the severity of the weather for that day. The severity check should be done every day in the game.

The following charts represent a mild climate. If the players are based in or travel to a different climate such as a desert or a mountain peak, add 10 to 20 on the 1d100 weather roll if the climate is hotter (10 to 20 because a desert is different to the tropics, both hotter but different to each other) and subtract 10 to 20 if the climate is colder.

The natural disaster result only applies if it is a natural 100 roll. If the characters are in a hotter climate and 10 is added to the weather roll, the natural disaster result only applies if a natural 100 was rolled, if 100 was achieved by adding to the roll, it is effectively 99, not 100.

Likewise, if the characters are in a colder climate and a 100 was rolled, the subtraction doesn’t count and the result is a natural disaster regardless.

Note; snow lasts much longer than any other weather condition, especially if the weather is cold. If snow was a result of a weather roll, and the following roll is cold, the snow remains on the ground.

\subsection{Spring}

\begin{center}
    \begin{xltabular}{\textwidth}{|c l r|} 
      \hline
      \textbf{Roll} & \textbf{Condition} & \textbf{Odds} \\
      \hline
      1 & Blizzard & 1\% \\ 
      2-4 & Snow & 3\% \\ 
      5-9 & Cold & 5\% \\
      10-19 & Storm & 10\% \\
      20-29 & Rain & 10\% \\
      30-49 & Overcast & 20\% \\
      50-69 & Mildly Cloudy & 20\% \\
      70-89 & Sunshine & 20\% \\
      90-99 & Hot & 10\% \\
      100 & Natural Disaster & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Summer}

\begin{center}
    \begin{xltabular}{\textwidth}{|c l r|} 
      \hline
      \textbf{Roll} & \textbf{Condition} & \textbf{Odds} \\
      \hline
      1 & Blizzard & 1\% \\ 
      2 & Snow & 1\% \\ 
      3-7 & Cold & 5\% \\
      8-12 & Storm & 5\% \\
      13-22 & Rain & 10\% \\
      23-32 & Overcast & 10\% \\
      33-52 & Mildly Cloudy & 20\% \\
      53-72 & Sunshine & 20\% \\
      73-99 & Hot & 27\% \\
      100 & Natural Disaster & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Autumn}

\begin{center}
    \begin{xltabular}{\textwidth}{|c l r|} 
      \hline
      \textbf{Roll} & \textbf{Condition} & \textbf{Odds} \\
      \hline
      1-2 & Blizzard & 2\% \\ 
      3-7 & Snow & 5\% \\ 
      8-17 & Cold & 10\% \\
      18-27 & Storm & 10\% \\
      28-47 & Rain & 20\% \\
      48-67 & Overcast & 20\% \\
      68-84 & Mildly Cloudy & 17\% \\
      85-94 & Sunshine & 10\% \\
      95-99 & Hot & 5\% \\
      100 & Natural Disaster & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Winter}

\begin{center}
    \begin{xltabular}{\textwidth}{|c l r|} 
      \hline
      \textbf{Roll} & \textbf{Condition} & \textbf{Odds} \\
      \hline
      1-5 & Blizzard & 5\% \\ 
      6-34 & Snow & 29\% \\ 
      34-74 & Cold & 40\% \\
      75-76 & Storm & 2\% \\
      77-81 & Rain & 5\% \\
      82-86 & Overcast & 5\% \\
      87-91 & Mildly Cloudy & 5\% \\
      92-97 & Sunshine & 6\% \\
      98-99 & Hot & 2\% \\
      100 & Natural Disaster & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Natural Disaster}

This will always depend on the climate and season. You may select the appropriate natural disaster and then roll how severe it is

\begin{itemize}
    \item Hurricane
    \item Tsunami
    \item Tornado
    \item Fire
    \item Earthquake
    \item Volcano
    \item Dust storm
\end{itemize}

\section{Open Land Encounters}

These random encounter rolls should be done every 1 to 10 kilometres

\subsection{Daytime Country Random Encounters}

\begin{center}
    \begin{xltabular}{\textwidth}{|c l r|} 
      \hline
      \textbf{Roll} & \textbf{Encounter} & \textbf{Odds} \\
      \hline
      1-16  & Nothing           & 16\% \\ 
      17-26 & Patrol            & 10\% \\ 
      27-38 & Peasant/farmer    & 12\% \\
      39-45 & Trade caravan     & 7\% \\
      46-52 & Noble Lord        & 7\% \\
      53-61 & Traveller         & 9\% \\
      62-67 & Adventurer        & 6\% \\
      68-73 & Wolf attack       & 6\% \\
      74-76 & Dire wolf attack  & 3\% \\
      77-85 & Bandit attack     & 9\% \\
      86-88 & Goblin attack     & 3\% \\
      89-90 & Ogre              & 2\% \\
      91-92 & Troll (near water)& 2\% \\
      93-94 & Orc raiding party & 2\% \\
      95    & Giant             & 1\% \\
      96-97 & Harpy attack      & 2\% \\
      98    & Witch ritual      & 1\% \\
      99    & Griffon           & 1\% \\
      100   & Dragon            & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Nighttime Country Random Encounters}

\begin{center}
    \begin{xltabular}{\textwidth}{|c X r|} 
      \hline
      \textbf{Roll} & \textbf{Encounter} & \textbf{Odds} \\
      \hline
      1-3   & Nothing           & 3\% \\ 
      4-6   & Patrol            & 3\% \\ 
      7     & Peasant/farmer    & 1\% \\
      8-9   & Trade caravan     & 2\% \\
      10    & Noble Lord        & 1\% \\
      11    & Traveller         & 1\% \\
      12-13 & Adventurer        & 2\% \\
      14-22 & Wolf attack       & 9\% \\
      23-26 & Dire wolf attack  & 4\% \\
      27    & Bandit attack     & 1\% \\
      28-41 & Goblin attack     & 14\% \\
      42-47 & Ogre              & 6\% \\
      48-52 & Troll (near water)& 5\% \\
      53-66 & Orc raiding party & 14\% \\
      67-80 & Undead \newline 
        \hspace{.8em} $\bullet$ Zombie \newline
        \hspace{.8em} $\bullet$ Skeleton\newline
        \hspace{.8em} $\bullet$ Wight \newline
        \hspace{.8em} $\bullet$ Wraith & 14\% \\
      81-84 & Giant             & 4\% \\
      85-88 & Harpy attack      & 4\% \\
      89-93 & Witch ritual      & 5\% \\
      94-98 & Werewolf          & 5\% \\
      99    & Griffon           & 1\% \\
      100   & Dragon            & 1\% \\
      \hline
    \end{xltabular}
\end{center}

\section{City Encounters}

This random encounter roll should be done once every city block

\subsection{Daytime City Random Encounters}

\begin{center}
    \begin{xltabular}{\textwidth}{|c X r|} 
      \hline
      \textbf{Roll} & \textbf{Encounter} & \textbf{Odds} \\
      \hline
      1-12  & Nothing           & 12\% \\ 
      13-20 & City Guard        & 8\% \\ 
      21-27 & Noble             & 7\% \\
      28-34 & Traveller         & 7\% \\
      35-40 & Adventurer        & 6\% \\
      41-47 & Performer         & 7\% \\
      48-54 & Preacher          & 7\% \\
      55-63 & Trader            & 9\% \\
      64-68 & Open argument     & 5\% \\
      69-74 & Public execution  & 6\% \\
      75-79 & Festival          & 5\% \\
      80-84 & Wedding           & 5\% \\
      85-89 & Witness a mugging & 5\% \\
      90-94 & Someone attempts to steal from party member & 5\% \\
      95-98 & Robbery at knife point & 4\% \\
      99-100& Fire              & 2\% \\
      \hline
    \end{xltabular}
\end{center}

\subsection{Nighttime City Random Encounters}

\begin{center}
    \begin{xltabular}{\textwidth}{|c X r|} 
      \hline
      \textbf{Roll} & \textbf{Encounter} & \textbf{Odds} \\
      \hline
      1-15  & Nothing               & 15\% \\
      16-30 & City patrol           & 15\% \\
      31-34 & Noble                 & 4\% \\
      35-40 & Traveller             & 6\% \\
      41-46 & Adventurer            & 6\% \\
      47-50 & Trader                & 4\% \\
      51-70 & Someone getting attacked & 20\% \\
      71-85 & Thieves attack party  & 15\% \\
      86-95 & Gargoyle attack       & 10\% \\
      96-100& Vampire               & 5\% \\
      \hline
    \end{xltabular}
\end{center}