The hardcore armour rules make armour significantly more powerful and reflect reality better.

Armour levels are calculated by adding the armour ratings of each specific piece of armour being worn, which will then give the character their total armour level.

Armour also adds weight which affects combat. Half the player's total armour level (rounded down) is subtracted from their total combat roll.

A character can then reduce injury levels by the equivalent amount of levels they have in armour.

The preset injury reduction that armour provides is permanent and applies to every injury the player receives in every combat encounter, meaning the only way to injure an opponent in armour is to exceed their armour score, or bypass it.

In these circumstances, the most efficient way of defeating a person in armour is grappling them or tripping them and holding them down, just like in real life.

Having said this, the hardcore armour rules are only truly balanced in the game mechanics if the Narrator and Players are also employing the hardcore weapon modifiers. Without the hardcore weapon modifiers, it will be very difficult, to impossible, to injure an opponent with a significantly high armour rating (unless you immobilize them or employ other combat modifiers.) A level eight victory is VERY hard to achieve, which would be needed to inflict even a minor injury to an opponent with an armour score of 7.

Below are the armour categories and if applicable in the setting a character can wear two pieces of armour in the same category, such as wearing a gambeson under chainmail or a breastplate, which was most certainly done in history. When this happens the armour rating stack, which is not necessarily a good thing as a higher armour rating subtracts more from a character’s combat roll.

The ratings as given are a guide and the Narrators can change this as they wish to make armour more or less significant in their game.

\begin{center}
    \begin{tabular}{|m{.7\textwidth} c|} 
      \hline
      \textbf{Armour Categories} & \textbf{Armour Level} \\
      \hline
      Small Shield (buckler) & 1 \\
      Medium shield (Heater shield, targe, small viking shield) & 2 \\
      Large shield (Kite Shield, scutum, large viking shield) & 3 \\
      Helmet & 1 \\
      Upper body protection, breastplate and fauld, gambeson, chainmail hauberk, kevlar & 1 \\
      Arm protection, pauldrons, arms and gauntlets & 1 \\
      Leg protection, cuisses and greaves & 1 \\
      \hline
    \end{tabular}
\end{center}