Thank you for checking out Cogent Roleplay. We hope you love it!

We’re still improving the rules and have many features planned – and we’d LOVE your feedback and support!

\section{Feedback and Support}

\href{http://discord.gg/5MrsGukbTq}{JOIN OUR DISCORD COMMUNITY} – we listen carefully to your experiences, good and bad. We know the only way to make Cogent the best it can be, is to get as much feedback from as many people as possible, and we care what you think, and want to hear about your role-play experiences!

We have been sharing Cogent for free for years now, and plan to keep the Cogent core rules completely free. We’ve already invested hundreds of hours and thousands of dollars into realizing the potential of Cogent Roleplay. To fully realize this potential, any financial support you can offer will be incredibly helpful!

\begin{itemize}
    \item \href{https://www.patreon.com/tabletoptime/}{SUPPORT US ON PATREON}, and get access to early features, and even custom assets and cameos!
    \item OR if you don’t want to commit to monthly support, a \href{https://www.buymeacoffee.com/tabletoptime}{ONE-TIME DONATION} would be incredible!
\end{itemize}

%[Error!] This is now untrue.
%Please note 100\% of donations and contributions go directly to assets, server costs/upgrades, Cogent app/website development, editing etc. We (Jazza and Shad) don’t need it, we're investing in our own! Your contribution will directly help Cogent Roleplay realize its potential!

Please note 100\% of donations and contributions go directly to TableTopTime, the original creators of Cogent. \textbf{They do not support the development of Cogent moving forward. That is a community effort.}

That said – the most important thing to us is that Cogent works for you, and you love playing it! If you do, the best support you can offer is to spread the word! If you make content of any kind, consider reviewing or paying for our system. If you have a local hobby shop or creative friends, see if they’d be open to trying it out with you! The more people that know about and play Cogent Roleplay, the stronger it will become!

Thank you for joining us on our quest – we wish you all the best in the many adventures ahead of you!

Many thanks,

Jazza and Shad